---
title: "Assignment_1"
output:
  pdf_document: default
  word_document: default
date: "2025-09-05"
---

## Task-1: Getting the dataset from web source

I downloaded the publicly accessible CNG stations' location data from the website of Department of Energy

Reference:
Office of Energy Efficiency and Renewable Energy. (n.d.). Natural Gas Fueling Station Locations.US Department of Energy. Retrieved from https://afdc.energy.gov/fuels/natural-gas-locations#/find/nearest?fuel=CNG


## Task-2: Importing the dataset dowloaded previously

```{r}
library(readr)
CNG_stations <- read_csv(".\\CNG_stations.csv")
View(CNG_stations)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Task-3: Descriptive analysis

```{r}
library(tidyverse)
#using quantitative variable
summary(CNG_stations$`CNG Total Compression Capacity`)
#state-wise number of cng stations
CNG_stations %>% group_by(`State`) %>% summarise(tot_station = n())

#using categorical variable
summary(CNG_stations$`Facility Type`)
#number of cng stations near different facilities
facility_stations <- CNG_stations %>% group_by(`Facility Type`) %>% summarise(station_no = n())
facility_stations
save(facility_stations,file=".\\facility_stations.csv")
```

## Task-4: Transforming variable

```{r}
is.logical(CNG_stations$`Restricted Access`)
#converting logical var to numeric
as.numeric(CNG_stations$`Restricted Access`)
```

## Task-5: Plotting

```{r}
library(ggplot2)
#histogram
hist(CNG_stations$`CNG Total Compression Capacity`)

#bar chart
ggplot(CNG_stations, aes(x = `Owner Type Code`)) +
  geom_bar(fill = "tomato") +
  labs(title = "Total number of stations by owner type", x = "Owner Category", y = "Count")

#scatter plot
ggplot(CNG_stations, aes(x = `CNG Dispenser Num`, y = `CNG Storage Capacity`)) +
  geom_point(color = "blue", size = 1.5) +
  theme_classic()+
  labs(title = "CNG storage capacity with dispenser numbers in stations")
```






